---
title: "Foundations of Data Science Individual Project"
author: "Matthias Bartolo"
date: "21/12/2021"
output: pdf_document
---

# Task 1: Queen's Gambit

## i)To find the probability that Beth wins the first game
The probability that Beth faces a beginner player is:\newline
__P(Beginner)__=$\frac{1}{3}$

The probability that Beth faces an intermediate player is:\newline
__P(Intermediate)__=$\frac{1}{3}$

The probability that Beth faces a master player is:\newline
__P(Master)__=$\frac{1}{3}$

The probability that Beth wins against  a beginner player is:\newline
__P(WinBeginner)__=$\frac{4}{5}$

The probability that Beth wins against  an intermediate player is:\newline
__P(WinIntermediate)__=$\frac{2}{5}$

The probability that Beth wins against  a master player is:\newline
__P(WinMaster)__=$\frac{1}{5}$
\newline
\newline
Therefore the Probability that Beth wins the first game is:\newline
\newline
__P(Beth Wins)__=P(Beginner)\*P(WinBeginner)+P(Intermediate)\*P(WinIntermediate)+P(Master)\*P(WinMaster)\newline\newline
__P(Beth Wins)__=$(\frac{1}{3} * \frac{4}{5}) + (\frac{1}{3} * \frac{2}{5}) + (\frac{1}{3} * \frac{1}{5}) = \frac{7}{15}$\newline\newline\newline\newline
__Therefore: P(Beth Wins)__=$\frac{7}{15}$\newline

\newpage
## ii)To find the probability that Beth wins the second game, if she won the first game
__Assuming the opponent doesn't change__\newline\newline
__P(Beth Wins Second game)__=\newline
P(Beginner)\*P(WinBeginner)\*P(WinBeginner)+\newline
P(Intermediate)\*P(WinIntermediate)\*P(WinIntermediate)+\newline
P(Master)\*P(WinMaster)\*P(WinMaster)\newline\newline
__We get:__\newline
__P(Beth Wins Second game)__=$(\frac{1}{3} * \frac{4}{5} * \frac{4}{5}) + (\frac{1}{3} * \frac{2}{5} * \frac{2}{5}) + (\frac{1}{3} * \frac{1}{5} * \frac{1}{5}) = \frac{7}{25}$\newline\newline

__Therefore: P(Beth Wins Second game)__=$\frac{7}{25}$\newline\newline\newline\newline\newline\newline


## iii)Independent and Conditionally Independent Outcomes
### Independent outcomes
Independent events describe situations when an event does not affect the probability of the other event occurring, as both events are disjoint events.

### Conditionally Independent outcomes
Conditionally Independent events describe situations when an event which is independent from the other event, will affect the other event based on a external parameter. In general taking events A and B which can be either dependent or independent on each other, they would be independent from each other when event C occurs (C being the external parameter).

For example, if the opponent is a beginner, then the probability of winning the first game is 4/5, and the probability of winning the second game is also 4/5 since the opponent doesn't change. In case the opponent changes, the probability of winning the first game remains the same whilst the probability of winning the second game depends on the opponent's skill level.


### Assumption
In this case I believe that the outcomes of the games are conditionally independent. Assuming that the opponent remains the same the probability of Beth winning each game is the same, as the probability of each game depends on the opponent's skill level. 

\newpage
# Task 2: Conditionitis
## i)To find the probability that a man has the disease given that he tested positive
__Let:__ \newline
1. __T   __ be the event that a man has a positive test\newline
2. __T^C^__ be the event that a man has a negative test\newline
3. __D   __  be the event that a man has actually the disease\newline
4. __D^C^__ be the event that a man does not have the disease\newline


The probability that a man test positives given he has the disease:\newline
__P(T|D)__=$\frac{9}{10}$

The probability that a man test negative given he has the disease:\newline
__P(T^C^|D)__=$\frac{1}{10}$

The probability that a man test positive given he doesn't have the disease:\newline
__P(T|D^C^)__=$\frac{1}{100}$

The probability that a man test negative given he doesn't have the disease:\newline
__P(T^C^|D^C^)__=$\frac{99}{100}$

The probability that a man has the disease:\newline
__P(D)__=$\frac{5}{100}$

The probability that a man doesn't have the disease:\newline
__P(D^C^)__=$\frac{95}{100}$\newline\newline

To obtain the following Probability Tree, a combination of the subsequent functions were used: geom_point, geom_segment as well as geom_text. Points were first plotted with the geom_point, then through the geom_segment the points were connected with a line. Furthermore, geom_text was also used to issue the respective text to the relevant lines.The following functions were also used in order to obtain a white background for the probability tree: theme_void(), theme(), xlab(NULL), ylab(NULL),scale_fill_viridis_c() and theme(). Finally, ggtitle() was used to issue a title to the plot.

```{r Probabilitytree, echo=FALSE}
library(ggplot2)
xlabel<-1
ylabel<-2


#Plotting the graph via ggplot
ggplot()+xlim(0,2.5)+ylim(-4,4)+
  ggtitle("Probability tree")+
    theme_void()+
  theme(plot.title = element_text(hjust = 0.5))+
  xlab(NULL)+ylab(NULL)+
  geom_tile() +
  scale_fill_viridis_c() + 
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank())+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())+
  geom_point(mapping=aes(x=xlabel-1,y=ylabel-1.5),color="black")+
  geom_point(mapping=aes(x=xlabel,y=ylabel),color="black")+
  geom_point(mapping=aes(x=xlabel,y=ylabel-3),color="black")+
  geom_segment(aes(x=xlabel-1,xend=xlabel,y=ylabel-1.5,yend=ylabel),color="black")+
  geom_text(aes(label="P(!D)",x=(xlabel)/2,y=(ylabel-2.5)/2),hjust=0, vjust=0,color="blue")+
  geom_text(aes(label="95/100",x=(xlabel)/2,y=(ylabel-2.5)/2),hjust=-0.5, vjust=3,color="purple")+
  geom_segment(aes(x=xlabel-1,xend=xlabel,y=ylabel-1.5,yend=ylabel-3),color="black")+
  geom_text(aes(label="P(D)",x=(xlabel)/2,y=(ylabel+1.25)/2),hjust=0, vjust=0,color="blue")+
  geom_text(aes(label="5/100",x=(xlabel)/2,y=(ylabel)-0.5),hjust=-0.5, vjust=2,color="purple")+
  geom_point(mapping=aes(x=xlabel+1,y=ylabel+1),color="black")+
  geom_text(aes(label="9/200",x=(xlabel+1),y=(ylabel+1)),hjust=0, vjust=2,color="purple")+
  geom_point(mapping=aes(x=xlabel+1,y=ylabel-1),color="black")+
  geom_text(aes(label="1/200",x=(xlabel+1),y=(ylabel-1)),hjust=0, vjust=2,color="purple")+
  geom_segment(aes(x=xlabel,xend=xlabel+1,y=ylabel,yend=ylabel+1),color="black")+
  geom_text(aes(label="P(T|D)",x=(xlabel)+0.5,y=(ylabel+0.8)),hjust=0, vjust=0,color="blue")+
  geom_text(aes(label="90/100",x=(xlabel)+0.5,y=(ylabel+0.8)),hjust=-0.5, vjust=2.25,color="purple")+
  geom_segment(aes(x=xlabel,xend=xlabel+1,y=ylabel,yend=ylabel-1),color="black")+
  geom_text(aes(label="P(T|!D)",x=(xlabel)+0.5,y=(ylabel-0.4)),hjust=0, vjust=0,color="blue")+
  geom_text(aes(label="10/100",x=(xlabel)+0.5,y=(ylabel-0.4)),hjust=-0.5, vjust=2.75,color="purple")+
  geom_point(mapping=aes(x=xlabel+1,y=ylabel-2),color="black")+
  geom_text(aes(label="19/2000",x=(xlabel+1),y=(ylabel-2)),hjust=0, vjust=2,color="purple")+
  geom_point(mapping=aes(x=xlabel+1,y=ylabel-4),color="black")+
  geom_text(aes(label="1881/2000",x=(xlabel+1),y=(ylabel-4)),hjust=0, vjust=2,color="purple")+
  geom_segment(aes(x=xlabel,xend=xlabel+1,y=ylabel-3,yend=ylabel-2),color="black")+
  geom_text(aes(label="P(T|D)",x=(xlabel)+0.5,y=(ylabel-2.2)),hjust=0, vjust=0,color="blue")+
  geom_text(aes(label="90/100",x=(xlabel)+0.5,y=(ylabel-2.2)),hjust=-0.5, vjust=2.25,color="purple")+
  geom_segment(aes(x=xlabel,xend=xlabel+1,y=ylabel-3,yend=ylabel-4),color="black")+
  geom_text(aes(label="P(T|!D)",x=(xlabel)+0.5,y=(ylabel-3.4)),hjust=0, vjust=0,color="blue")+
  geom_text(aes(label="10/100",x=(xlabel)+0.5,y=(ylabel-3.4)),hjust=-0.5, vjust=2.75,color="purple")
```
__Required to find: P(D|T)__\newline\newline
__Formula:__\newline\newline
__P(D|T)__=$\frac{P(D)P(T|D)}{P(T)}$\newline\newline\newline

__where:__\newline
__P(T)__=P($D \cap T$)+P($D^{C} \cap T$)\newline\newline
__P(T)__=P(D)\*P(T|D)+P(D^C^)\*P(T|D^C^)\newline\newline
__P(T)__=$(\frac{5}{100} * \frac{90}{100}) + (\frac{95}{100} * \frac{1}{100}) = \frac{109}{2000}$\newline\newline\newline
__Putting into formula:__\newline \newline __P(D|T)__=$\frac{P(D)P(T|D)}{P(T)}$ \newline\newline
__We get:__\newline$(\frac{5}{100} * \frac{90}{100}) / \frac{109}{2000}= \frac{90}{109}$\newline\newline\newline
__Therefore: P(D|T)__=$\frac{90}{109}$

## ii)To find the probability that a man has the disease given that he has a negative test

__Required to find: P(D|T^C^)__\newline\newline
__Formula:__\newline\newline
__P(D|T^C^)__=$\frac{P(D)P(T^{C}|D)}{P(T^{C})}$\newline\newline\newline

__where:__\newline
__P(T^C^)__=P($D \cap T^{C}$)+P($D^{C} \cap T^{C}$)\newline\newline
__P(T^C^)__=P(D)\*P(T^C^|D)+P(D^C^)\*P(T^C^|D^C^)\newline\newline
__P(T^C^)__=$(\frac{5}{100} * \frac{10}{100}) + (\frac{95}{100} * \frac{99}{100}) = \frac{1891}{2000}$\newline\newline\newline
__Putting into formula:__\newline \newline __P(D|T^C^)__=$\frac{P(D)P(T^{C}|D)}{P(T^{C})}$ \newline\newline
__We get:__\newline$(\frac{10}{100} * \frac{90}{100}) / \frac{1891}{2000}= \frac{10}{1891}$\newline\newline\newline
__Therefore: P(D|T^C^)__=$\frac{10}{1891}$


\newpage
# Task 3: Bottled Water
## i)Null and Alternative Hypothesis

__Null Hypothesis:__ $H_{0}$: $\theta$ = 240ml \newline The machine continues to dispense water normally, since the amount of water dispensed is equal to 240ml, and thus the machine does not need to be stopped.\newline\newline
__Alternate Hypothesis:__ $H_{1}$: $\theta$ $\neq$ 240ml \newline The machine should be stopped and production waits for repairs, since the amount of water is not equal to 240ml. \newline\newline

## ii)Distribution Diagram

To obtain the following Distribution Diagram, a combination of the subsequent functions were used: geom_point, geom_segment, geom_line, ylab, xlab, theme_bw, theme and scale_x_continuous.\newline\newline

```{r Distribution, echo=FALSE}
library(ggplot2)
#Variance which was given
variance<-0.88
#Creating a vector which represents the x-axis
xval <- seq(-3, 3, by = .001)
xlab2 <- seq(-3, 3, by = 1)
#Creating a vector which calculates the values in the y-axis through dnorm function
yval <- dnorm(xval, mean = 0, sd = sqrt(variance))
#Creating a data frame and combining both vectors
data<-data.frame(xval,yval)

#Plotting the graph and respective critical value lines through functions seen below
plotd<-ggplot(data,aes(x=xval,y=yval))+
  ggtitle("Distribution Diagram")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))+
  xlab("z-scores")+
  theme(axis.text.x = element_text(angle = 360, vjust = 0.5)) +
  scale_x_continuous(labels=as.character(xlab2),breaks=xlab2)+
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())+
  ylab("f")+
  geom_point(color="blue")+
  geom_line()+
  geom_point(mapping=aes(x=data[which(xval==0),1],y=data[which(xval==0),2]),color="blue")+
  geom_point(mapping=aes(x=0,y=0),color="purple")+
  geom_segment(aes(x=0,xend=data[which(xval==0),1],y=0,yend=data[which(xval==0),2]),size=2.5,color="purple")+
  geom_point(mapping=aes(x=data[which(xval==2.58),1],y=data[which(xval==2.58),2]),color="blue")+
  geom_point(mapping=aes(x=2.58,y=0),color="orange")+
  geom_segment(aes(x=2.58,xend=data[which(xval==2.58),1],y=0,yend=data[which(xval==2.58),2]),size=2.5,color="orange")+
  geom_point(mapping=aes(x=data[which(xval==-2.58),1],y=data[which(xval==-2.58),2]),color="blue")+
  geom_point(mapping=aes(x=-2.58,y=0),color="green")+
  geom_segment(aes(x=-2.58,xend=data[which(xval==-2.58),1],y=0,yend=data[which(xval==-2.58),2]),size=2.5,color="green")
  
  plotd
```
\newpage

__Given:__\newline
__1.__ Sample Mean: $\bar{x}=234ml$\newline
__2.__ Population Mean: $\mu=240ml$\newline
__3.__ Sample Size: $N=35$\newline
__4.__ Variance: $\sigma^{2}=0.88$\newline
__5.__ Significance: $\alpha = 0.99$\newline\newline\newline

__To obtain the Distribution diagram;__\newline

1.First the standard deviation was found by obtaining the square root of the variance (0.88). i.e\space\space$\sigma=0.93$\newline\newline
2.The level of significance (0.99) was subtracted from 1 and 0.01 was obtained. The latter (0.01) was then divided by 2, since a two-tailed test was conducted and 0.005 was achieved. The value (0.005) was subtracted from 1 and the result was used in the z-table to find the critical values (green/orange values in graph).\newline\newline
3.The dnorm function was utilized to populate the y values to plot a normal distribution graph.\newline\newline
4.Furthermore, the rejection regions are the regions which are smaller than -2.58 & larger than 2.58\newline(-2.58 and 2.58 being the critical values).\newline\newline\newline
The purple line signifies the mean which is 240ml.\newline\newline\newline Link to z-table used:\newline https://www.statisticshowto.com/tables/z-table/

\newpage

## iii)Sample parameters and test-statistics

__Finding the Test-Statistic:__\newline\newline __Formula used:__\newline
__z__=$\frac{\bar x - \mu}{\sigma_{\bar x}}$\newline\newline\newline
__Where:__\newline
__1.__ z: Standard Score\newline
__2.__ $\bar x$: Observed value\newline
__3.__ $\mu$: Mean of sample\newline
__4.__ $\sigma_{\bar x}$: Standard Error of means\newline\newline\newline

__Finding the Standard Error of means:__\newline\newline
$\sigma_{\bar x}=\frac{\sigma_{x}}{\sqrt{N} }= \frac{0.93}{\sqrt{35}}$\newline\newline\newline
$\sigma_{\bar x}=0.159$\newline\newline\newline
__Substituting in the Formula:__\newline\newline
__z__=$\frac{234 - 240}{0.159}$\newline\newline\newline
__Therefore Test Statistic: z__ = -37.84\newline

## iv)Comparing the test-statistic with the critical values

When comparing the test-statistic of -37.84 with the left-most critical value of -2.58 (green value in graph), it was noted that the test-statistic was smaller than the critical value. This would imply that the test-statistic resides in the left rejection region.\newline

## v)Conclusion

In conclusion, due to the test-statistic being smaller than the critical value, the null hypothesis is rejected, since the conclusion has a significance of more than 99%. Thus, The alternate hypothesis is utilized where the machine should be stopped and production waits for repairs.
